---
title: "K Means Shiny App"
author: "Evan Johnson"
date: '`r Sys.Date()`'
output: html_document
---
```{r setiup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
```

## K means example {.tabset}

This is a break down for the k-means example shiny app into easy to understand steps, adding a few shiny elements at a time. You can follow along and build the app using the template. Is is adapted from the [Kmeans example code from Joe Cheng at Rstudio](https://shiny.rstudio.com/gallery/kmeans-example.html)

### Step 00: Template
```{r, template, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      
    )
  )
)

server <- function(input, output, session) {

}

shinyApp(ui = ui, server = server)
```

### Step 01: Title Panel
```{r, add title panel, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      
    )
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui = ui, server = server)
```

### Step 02: UI inputs
```{r, add ui inputs, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      
    )
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui = ui, server = server)
```

### Step 03: UI outputs
```{r, add ui outputs, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui = ui, server = server)
```

### Step 04: Output plot
```{r, add output plot, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  output$plot1 <- renderPlot({
    plot(iris[,c("Sepal.Length","Sepal.Width")])
  })
}

shinyApp(ui = ui, server = server)
```

### Step 05: Reactive Axes
```{r, add reactive axes, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  output$plot1 <- renderPlot({
    plot(iris[,c(input$xcol,input$ycol)])
  })
}

shinyApp(ui = ui, server = server)
```

### Step 06: K means
```{r, add K means clustering, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {

  output$plot1 <- renderPlot({
    plot(iris[,c(input$xcol,input$ycol)],
         col = kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)$cluster)
  })
}

shinyApp(ui = ui, server = server)
```

### Step 07: Cluster centers
```{r, add cluster centers, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  
  output$plot1 <- renderPlot({
    plot(iris[,c(input$xcol,input$ycol)],
         col = kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)$cluster)
    points(kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)$centers, pch = 4, cex = 4, lwd = 4)
  })
}

shinyApp(ui = ui, server = server)
```

### Step 08: Save Kmeans
```{r, save k means result as reactive, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  
  clusters <- reactive({
    kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)
  })
  
  output$plot1 <- renderPlot({
    plot(iris[,c(input$xcol,input$ycol)],
         col = clusters()$cluster)
    points(clusters()$centers, pch = 4, cex = 4, lwd = 4)
  })
}

shinyApp(ui = ui, server = server)
```

### Step 09: Final App
```{r, final app, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  
  selectedData <- reactive({
    iris[, c(input$xcol, input$ycol)]
  })
  
  clusters <- reactive({
    kmeans(selectedData(), input$clusters)
  })
  
  output$plot1 <- renderPlot({
    palette(c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
              "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))
    
    par(mar = c(5.1, 4.1, 0, 1))
    plot(selectedData(),
         col = clusters()$cluster,
         pch = 20, cex = 3)
    points(clusters()$centers, pch = 4, cex = 4, lwd = 4)
  })
}

shinyApp(ui = ui, server = server)
```